---
title: "Followup 1 - GLM"
author: "Subbiah"
date: "18 June 2017"
output: html_document
---
Coding a categorical predictor is not uncommon in any GLM.**Any arbitrary coding will not affect the estimates.** <br>This notes provides an illustration with Logistic regression in classical as well as Bayesian procedure<br><br>
**CLASSICAL INFERENCE**
```{r}
#data
dat_LR10RES<-list(
sex=c(1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 0, 1, 0, 0, 0, 1, 1, 1, 0, 0, 1,1, 0, 1, 1, 1, 0, 0, 0, 1, 1, 0, 0, 1, 1, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 1, 0, 1, 1, 1, 0, 0, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 0, 0, 1, 1, 1, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 0, 0, 1, 1, 1, 0, 1, 0, 1, 1, 1, 0, 1, 1, 1, 1, 1),
age= c(69, 57, 61, 60, 69, 74, 63, 68, 64, 53, 60, 58, 79, 56, 53, 74, 56, 76, 72,56, 66, 52, 77, 70, 69, 76, 72, 53, 69, 59, 73, 77, 55, 77, 68, 62, 56, 68, 70, 60,65, 55, 64, 75, 60, 67, 61, 69, 75, 68, 72, 71, 54, 52, 54, 50, 75, 59, 65, 60, 60, 57, 51, 51, 63, 57, 80, 52, 65, 72, 80, 73, 76, 79, 66, 51, 76, 75, 66, 75, 78, 70, 67, 51, 70, 71, 71, 74, 74, 60, 58, 55, 61, 65, 52, 68, 75, 52, 53, 70),
frac=c(1, 1, 1, 0, 1, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0,1, 0, 1, 1, 0, 1, 1, 1, 0, 1, 1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1,1, 1, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 1, 1, 1, 0, 0, 1, 1, 0, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1,1, 0, 1, 1, 0, 0, 1, 0, 0, 1),n=100)
#Both response and predictor are coded
frac = factor(dat_LR10RES$frac, levels=c(0,1))
sex=factor(dat_LR10RES$sex, levels=c(0,1))
summary(glm(frac ~ dat_LR10RES$age +sex, family = binomial))
#It is possible to observe that smaller code is considered as reference category. If coding is changed so that existing 1 becomes smaller
sex_R=0
for(i in 1:100)
{if(sex[i]==0) sex_R[i]=2 else sex_R[i]=1}
sex_R=factor(sex_R)
summary(glm(frac ~ dat_LR10RES$age +sex_R, family = binomial))
```
you could see the changes in the output<br><br>
"sex1" is the variable name in the output when the smaller coding (0) is the reference category.<br>
"sex_R2" is the variable name in the output when the smaller coding (1) is the reference category. In this case 0 is recoded.<br><br>
The estimates for the categorical predictor is having opposite signs when the reference category is changed. Also to observe that no other changes in the estimates and model fit constants, number of iterations.<br> This is the effect of such reversal coding and hence this may not be necessary in practice, if one needs to reverse the comparison only the sign change is needed and subsequent interpretation should be proper<br><br>
*Finally one more illustation of this. Now let us recode 0 as -1 and 1 as 1 so that -1 (earlier 0) is reference category*
```{r}
sex_RR=0
for(i in 1:100)
{if(sex[i]==0) sex_RR[i]=-1 else sex_RR[i]=1}
sex_RR=factor(sex_RR)
summary(glm(frac ~ dat_LR10RES$age +sex_RR, family = binomial))
```
You could see the similarity<br>
As far as response variable is concerned it is a practice to code the success (desired) event of the response variable as 1 and other with smaller one<br><br>
***BAYESIAN INFERENCE <br>***
**Before run the code<br>**
\Bullet 1 Make sure that bugs.directory is set properly <br>
\Bullet 2 debug option is FALSE. Make it TRUE if needed
```{r}
require(R2WinBUGS)
#LR10RES Logistic regression model with TWO predictor - AGE metric and sex categorical
LR10RES<-function()
{
  for (i in 1:n) {
    # Likelihood function for each data point
    frac[i] ~ dbern(p[i])
  # S2[i]           <- equals(sex[i], 0) #<- zero coded is used in analysis; ref category:1
    S2[i]           <- equals(sex[i], 1) #<- one coded is used in analysis; ref category:0
    # Linear regression on logit
    logit(p[i]) <- alpha + b.sex*S2[i] + b.age*age[i]
  }
  alpha ~ dnorm(1,1.0E-4) # Prior for intercept
  b.sex ~ dnorm(0.0,1.0E-4) # Prior for slope of sex
  b.age ~ dnorm(0.5,1.0E-2) # Prior for slope of age
}
#Writing BUGS File 
LR10RES_FILE<- file.path(tempdir(), "LR10RES.bug")
write.model(LR10RES, LR10RES_FILE)
#file.show(LR10RES_FILE) #optional
dat_LR10RES<-list(
sex=c(1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 1, 1, 1, 1, 1, 0, 1, 0, 0, 0, 1, 1, 1, 0, 0, 1,1, 0, 1, 1, 1, 0, 0, 0, 1, 1, 0, 0, 1, 1, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 1, 0, 1, 1, 1, 0,0, 1, 1, 1, 1, 0, 0, 0, 1, 1, 1, 0, 0, 1, 1, 1, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 0, 0, 1, 1, 1,0, 1, 0, 1, 1, 1, 0, 1, 1, 1, 1, 1),
age= c(69, 57, 61, 60, 69, 74, 63, 68, 64, 53, 60, 58, 79, 56, 53, 74, 56, 76, 72,56, 66, 52, 77, 70, 69, 76, 72, 53, 69, 59, 73, 77, 55, 77, 68, 62, 56, 68, 70, 60,65, 55, 64, 75, 60, 67, 61, 69, 75, 68, 72, 71, 54, 52, 54, 50, 75, 59, 65, 60, 60, 57, 51, 51, 63, 57, 80, 52, 65, 72, 80, 73, 76, 79, 66, 51, 76, 75, 66, 75, 78, 70, 67, 51, 70, 71, 71, 74, 74, 60, 58, 55, 61, 65, 52, 68, 75, 52, 53, 70),
frac=c(1, 1, 1, 0, 1, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0,1, 0, 1, 1, 0, 1, 1, 1, 0, 1, 1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1,1, 1, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 1, 1, 1, 0, 0, 1, 1, 0, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1,1, 0, 1, 1, 0, 0, 1, 0, 0, 1),n=100)
frac = factor(dat_LR10RES$frac, levels=c(0,1))
sex=factor(dat_LR10RES$sex, levels=c(0,1))
ns=500#Number of simulations in BUGS
inits <- function()
{
  list(alpha=0, b.sex=1, b.age=1)
  
}
parameters <- c("alpha","b.sex","b.age") #no need to obtaine the estimates of theta - if model is desired
post_LR10RES.sim <- bugs(dat_LR10RES, inits, parameters, model.file=LR10RES_FILE,n.chains=1,n.thin=1, n.iter=ns,bugs.directory="D:/WinBUGS14/",debug =FALSE,digits=3)
print(post_LR10RES.sim,digits=3)
```
**Now change the reference category**
```{r}
LR10RES_R<-function()
{
  for (i in 1:n) {
    # Likelihood function for each data point
    frac[i] ~ dbern(p[i])
    S2[i]           <- equals(sex[i], 0) #<- zero coded is used in analysis; ref category:1
    #S2[i]           <- equals(sex[i], 1) #<- one coded is used in analysis; ref category:0
    # Linear regression on logit
    logit(p[i]) <- alpha + b.sex*S2[i] + b.age*age[i]
  }
  alpha ~ dnorm(1,1.0E-4) # Prior for intercept
  b.sex ~ dnorm(0.0,1.0E-4) # Prior for slope of sex
  b.age ~ dnorm(0.5,1.0E-2) # Prior for slope of age
}
LR10RES_R_FILE<- file.path(tempdir(), "LR10RES_R.bug")
write.model(LR10RES_R, LR10RES_R_FILE)
post_LR10RES_R.sim <- bugs(dat_LR10RES, inits, parameters, model.file=LR10RES_R_FILE,n.chains=1,n.thin=1, n.iter=ns,bugs.directory="D:/WinBUGS14/",debug =FALSE,digits=3)
print(post_LR10RES_R.sim,digits=3)
```
**Change the coding from zero to -1 and no change for code 1**
```{r}
dat_LR10RES_RR<-list(sex_RR=c(1, 1, 1, -1, 1, 1, -1, -1, -1, -1, 1, 1, 1, 1, 1, -1, 1, -1, -1, -1, 1, 1, 1, -1, -1, 1, 1, -1, 1, 1, 1, -1, -1, -1, 1, 1, -1, -1, 1, 1, -1, 1, -1, -1, -1, 1, -1, -1, -1, 1, -1, 1, -1, 1, 1, 1, -1, -1, 1, 1, 1, 1, -1, -1, -1, 1, 1, 1, -1, -1, 1, 1, 1, -1, -1, -1, 1, 1, -1, -1, -1, -1, 1, -1, -1, 1, 1, 1, -1, 1, -1, 1, 1, 1, -1, 1, 1, 1, 1, 1), age= c(69, 57, 61, 60, 69, 74, 63, 68, 64, 53, 60, 58, 79, 56, 53, 74, 56, 76, 72,56, 66, 52, 77, 70, 69, 76, 72, 53, 69, 59, 73, 77, 55, 77, 68, 62, 56, 68, 70, 60,65, 55, 64, 75, 60, 67, 61, 69, 75, 68, 72, 71, 54, 52, 54, 50, 75, 59, 65, 60, 60, 57, 51, 51, 63, 57, 80, 52, 65, 72, 80, 73, 76, 79, 66, 51, 76, 75, 66, 75, 78, 70, 67, 51, 70, 71, 71, 74, 74, 60, 58, 55, 61, 65, 52, 68, 75, 52, 53, 70), frac=c(1, 1, 1, 0, 1, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0,1, 0, 1, 1, 0, 1, 1, 1, 0, 1, 1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0, 0, 0, 0, 1, 1, 1,1, 1, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 1, 1, 1, 0, 0, 1, 1, 0, 1, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1,1, 0, 1, 1, 0, 0, 1, 0, 0, 1),n=100)
LR10RES_RR<-function()
{
  for (i in 1:n) {
    # Likelihood function for each data point
    frac[i] ~ dbern(p[i])
    S2[i]  <- equals(sex_RR[i], -1) # zero (now -1) coded is used in analysis; ref category:1
    # Linear regression on logit
    logit(p[i]) <- alpha + b.sex*S2[i] + b.age*age[i]
  }
  alpha ~ dnorm(1,1.0E-4) # Prior for intercept
  b.sex ~ dnorm(0.0,1.0E-4) # Prior for slope of sex
  b.age ~ dnorm(0.5,1.0E-2) # Prior for slope of age
}
LR10RES_RR_FILE<- file.path(tempdir(), "LR10RES_RR.bug")
write.model(LR10RES_RR, LR10RES_RR_FILE)
post_LR10RES_RR.sim <- bugs(dat_LR10RES_RR, inits, parameters, model.file=LR10RES_RR_FILE, n.chains=1, n.thin=1,n.iter=ns,bugs.directory="D:/WinBUGS14/",debug =FALSE,digits=3)
print(post_LR10RES_RR.sim,digits=3)
```
<b><i> Hope this exemplifies the effect of coding a predictor.